---
title: "Untitled"
author: "Aaron W"
date: "February 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(XML)
library(httr)

library(dplyr)
library(leaflet)
library(knitr)
```

# Map of 2016 Population of each US state

This leaflet page shows the population estimates of each US state in 2016 with colored bubbles.

## Loading and processing data for map

Before creating an interactive map, first data on the center latitude and longitude of each US state
is collected from [this website](https://inkplant.com/code/state-latitudes-longitudes).

```{r latlong, results="asis", cache=TRUE, message=FALSE}
url <- "https://inkplant.com/code/state-latitudes-longitudes"
html <- GET(url)
content <- content(html, as="text")
parsedHTML <- htmlParse(content, asText=TRUE)
latlong_table <- readHTMLTable(parsedHTML, header=TRUE)[[1]]
```

Then, the population data for each state is collected from the US census. The
two data sets are then merged.

```{r pop, cache=TRUE, message=FALSE, warning=FALSE}
url2 <- "http://www2.census.gov/programs-surveys/popest/datasets/2010-2016/national/totals/nst-est2016-alldata.csv"
download.file("url2", "nst-est2016-alldata.csv", method="curl")
census <- read.csv("nst-est2016-alldata.csv")

# subset data to only include states and population estimate for 2016
census <- census %>%
     filter(STATE > 0 & STATE < 57) %>%
     select(State = NAME, Population = POPESTIMATE2016)

# merge data from two data sets
combined <- merge(census, latlong_table, by="State")
```

## Map creation

Finally, each state is marked with a bubble corresponding to its population
in leaflet.

```{r leaflet}
# Add colors
set.seed(334)
combined <- mutate(combined, color = sample(c("red", "blue", "green"), 51, replace = TRUE))

# Add circles for population and center map
combined %>%
     leaflet() %>%
     addTiles() %>%
     addCircles(lng = ~Longitude, lat = ~Latitude,
                weight = 2, radius = sqrt(combined$Population) * 50, 
                color = ~color, popup = ~State) %>%
     setView(lng = -120, lat = 40, zoom = 3)

```

## Appendix

The combined data used in this page is shown below.

```{r table, results='asis'}

kable(combined)
```